Qualitative Probabilistic Programming

### 2015-02-25 paper/paper.tex
paper and clustering

  Here is an example of a Quipp program to perform clustering:

@@

    In the factor graph, there are 2 types of variables: real numbers
    and categorical values.
    In Quipp, there are more value types than this.  Specifically,
    in addition to real numbers and categorical types, tuples
    are also supported.
    To translate a Quipp program to a factor graph, we
    map each value to a list of variables, and create factors
    over all these variables.

@@

>>>>

  Quipp is implemented as a Python library, so Quipp programs are written as Python programs that have access
  to special functions.  Here is an example of a Quipp program to perform clustering:

@@

    In the factor graph, there are 2 types of variables: real numbers
    and categorical values.
    In Quipp, we have an additonal tuple type, which gets translated
    to multiple underlying variables in the factor graph.

@@

  \subsection{Clustering}
\begin{verbatim}
dim = 2
nclusters = 3

PointType = Vector(dim, Double)
ClusterType = Categorical(nclusters)

In this example, we cluster 2d points into 3 different clusters.  Given a cluster, the distribution for a point is some independent Gaussian distribution.  This is similar to fuzzy c-means clustering.

  \subsection{Factor Analysis}

\begin{verbatim}
num_components = 3
point_dim = 5

ComponentsType = Vector(num_components, Double)
PointType = Vector(point_dim, Double)

get_point = randFunction(ComponentsType, PointType)

def sample():
  components = [normal(0, 1) for i in range(num_components)]
  return (comenents, get_point(components))
\end{verbatim}

The factor analysis model is very similar to the clustering model.  The main difference is that we replace the categorical \texttt{ClusterType} type with a vector type.  This results in the model attempting to find each point as an affine function of a vector of standard normal values.

  \subsection{Hidden Markov model}
\begin{verbatim}
num_states = 5
num_observations = 3
sequence_length = 10

StateType = Categorical(num_states)
ObsType = Categorical(num_observations)

trans_fun = rand_function(StateType, StateType)
obs_fun = rand_function(StateType, ObsType)

def sample():
  states = [uniform(num_states)]
  for i in range(sequence_length - 1):
    states.append(trans_fun(states[-1]))
  return (states, [obs_fun(s) for s in states])
\end{verbatim}

In this example, we use the unknown function \texttt{trans\_fun} for state transitions and \texttt{obs\_fun} for observations.  This means that we will learn both the state transitions and the observation distribution.

\subsection{Latent Dirichlet Allocation}

\begin{verbatim}
n_classes = 10
n_words = 10000
n_words_per_document = 1000

ClassType = Categorical(n_classes)
WordType = Categorical(n_words)

class_to_word = rand_function(ClassType, WordType)

def sample():
    which_class = uniform_categorical(n_classes)
    words = [class_to_word(which_class) for i in range(n_words_per_document)]
    return (which_class, words)
\end{verbatim}
